### 2014-10-08 Studies/2011_San_Francisco/Report/San_Francisco_Fire.tex
San Francisco: Edits to content and figures in Assessing the Hazard section

% PDF metadata
\hypersetup{
  pdftitle={Simulation of a Fire in a Hillside Residential Structure - San Francisco, CA},
  pdfauthor={Overholt, Kristopher J., Weinschenk, Craig G., and Madrzykowski, Daniel},
  pdfsubject={The Fire Dynamics Simulator (FDS) fire model, which is developed and maintained by the National Institute of Standards and Technology (NIST), was used to provide insight into the dynamics of a fire that occurred on June 2, 2011, within a multi-level, single-family residential structure in San Francisco, CA, that resulted in the death of two firefighters. The inputs for the FDS simulations are documented in this report and are based on the fire scenario, including the building geometry, interior furnishings, and ventilation conditions. The fire started in the basement and resulted in ventilation limited (fuel rich) fire conditions in the basement area. The fire originated in the basement, and the interior stairwell acted as a chimney for hot gases in the basement to flow towards areas of lower pressure via the interior stairwell towards the front of the structure. After the rear basement windows began to fail, the combustion gases at elevated temperature and pressure in the basement flowed upwards towards lower pressure areas via the stairwell. The temperature of the gases in the interior stairwell was estimated to be in excess of 760~$^{\circ}$C (1400~$^{\circ}$F), which exceeds the Class~III exposure temperature of 260~$^{\circ}$C (500~$^{\circ}$F). Unfortunately, two E26 firefighters were located in the flow path between the basement and the doors on the front side of the structure. After a Mayday call, both firefighters were removed from the structure and immediate medical treatment was provided. The two firefighters were transported to the local medical center where the lieutenant was pronounced dead and the fire fighter/paramedic died two days later.},
  pdfkeywords={flow path; fire dynamics; LODD; fire modeling; FDS simulation}
}

@@

The fire originated in the basement, and the interior stairwell acted as a chimney for hot gases in the basement to flow towards areas of lower pressure via the interior stairwell towards the front of the structure. After the rear basement windows began to fail, the combustion gases at elevated temperature and pressure in the basement flowed upwards towards lower pressure areas via the stairwell. The temperature of the gases in the interior stairwell was estimated to be in excess of 760~$^{\circ}$C (1400~$^{\circ}$F), which exceeds the Class~III exposure temperature of 260~$^{\circ}$C (500~$^{\circ}$F). Unfortunately, two E26 firefighters were located in the flow path between the basement and the doors on the front side of the structure. After a Mayday call, both firefighters were removed from the structure and immediate medical treatment was provided. The two firefighters were transported to the local medical center where the lieutenant was pronounced dead and the fire fighter/paramedic died two days later.

@@

Figure~\ref{fig:smv_flow_path_top_1} shows a top view of the temperature contours (at a height of 1.2~m above the floor) on the first floor, 2~s before the first rear basement window failed. At this time, the temperature of the gases in the interior stairwell exceeded 200~$^{\circ}$C (400~$^{\circ}$F).

\begin{figure}[!ht]
\includegraphics[width=4.3in]{../Figures/SMV_Top_Temp_340_s}
\input{../Figures/colorbar_metric_temp}
\input{../Figures/colorbar_std_temp}
\caption[Simulated temperature contours on the first floor before basement window failure.]
{Top view of simulated temperature contours on the first floor, 1.2~m above the floor (mid-level height), 2~s before the first rear basement window failure.}
\label{fig:smv_flow_path_top_1}
\end{figure}

Figure~\ref{fig:smv_flow_path_top_2} shows a top view of the temperature contours (at a height of 1.2~m above the floor) on the first floor, 60~s and 120~s after the first rear basement window failed. At a time of 60~s after the first rear basement window failed, the temperature of the gases in the interior stairwell exceeded 540~$^{\circ}$C (1000~$^{\circ}$F), which exceeds the Class~III exposure temperature of 260~$^{\circ}$C (500~$^{\circ}$F). As more rear basement windows failed and provided ambient air to the growing basement fire, the temperatures in the interior stairwell continued to rise. At a time of 120~s after the first rear basement window failed, the temperature of the gases in the interior stairwell exceeded 760~$^{\circ}$C (1400~$^{\circ}$F), which exceeds the Class~III exposure temperature of 260~$^{\circ}$C (500~$^{\circ}$F).

@@

\begin{figure}[!ht]
\includegraphics[width=4.3in]{../Figures/SMV_Top_Temp_402_s}
\input{../Figures/colorbar_metric_temp}
\input{../Figures/colorbar_std_temp}
\includegraphics[width=4.3in]{../Figures/SMV_Top_Temp_462_s}
\input{../Figures/colorbar_metric_temp}
\input{../Figures/colorbar_std_temp}
\caption[Simulated temperature contours on the first floor after basement window failure.]
{Top view of simulated temperature contours on the first floor, 1.2~m above the floor (mid-level height), 60~s (top) and 120~s (bottom) after the first rear basement window failure.}
\label{fig:smv_flow_path_top_2}
\end{figure}

Based on these results and the simulation results shown in Sections~\ref{sec:velocity} and \ref{sec:temperature}, after the rear basement windows failed, the simulated flow path temperatures in the interior stairwell were in excess of 760~$^{\circ}$C (1400~$^{\circ}$F) and the flow velocities were approximately 9~m/s (20 mph). The conditions in the interior stairwell changed from tenable to high-hazard very rapidly following the rear basement window failure. Within the structure, the hot gases and smoke were moving  along the flow path from the basement towards the doors located on the front side of the structure via the interior stairwell. The simulated temperatures are consistent with the post-incident conditions that were documented in the interior stairwell.

@@

The fire originated in the basement, and the interior stairwell acted as a chimney for hot gases in the basement to flow towards areas of lower pressure via the interior stairwell towards the front of the structure. After the rear basement windows failed, a flow path was established between the basement living room area and the doors located on the front side of the structure (the front door and the garage door). The rear basement window failures resulted in a rapid change in the conditions within the flow path. After the rear basement windows began to fail, the combustion gases at elevated temperature and pressure in the basement flowed upwards towards lower pressure areas via the stairwell. In the interior stairwell, the flow velocities were approximately 9~m/s (20 mph) and the temperature of the gases was estimated to be in excess of 760~$^{\circ}$C (1400~$^{\circ}$F), which exceeds the Class~III exposure temperature of 260~$^{\circ}$C (500~$^{\circ}$F).

>>>>

% PDF metadata
\hypersetup{
  pdftitle={Simulation of a Fire in a Hillside Residential Structure - San Francisco, CA},
  pdfauthor={Overholt, Kristopher J., Weinschenk, Craig G., and Madrzykowski, Daniel},
  pdfsubject={The Fire Dynamics Simulator (FDS) fire model, which is developed and maintained by the National Institute of Standards and Technology (NIST), was used to provide insight into the dynamics of a fire that occurred on June 2, 2011, within a multi-level, single-family residential structure in San Francisco, CA, that resulted in the death of two firefighters. The inputs for the FDS simulations are documented in this report and are based on the fire scenario, including the building geometry, interior furnishings, and ventilation conditions. The fire started in the basement and resulted in ventilation limited (fuel rich) fire conditions in the basement area. The fire originated in the basement, and the interior stairwell acted as a chimney for hot gases in the basement to flow towards areas of lower pressure via the interior stairwell towards the front of the structure. After the rear basement windows began to fail, the combustion gases at elevated temperature and pressure in the basement flowed upwards towards lower pressure areas via the stairwell. The temperature of the gases in the interior stairwell was estimated to be in excess of 700~$^{\circ}$C (1300~$^{\circ}$F), which exceeds the Class~III exposure temperature of 260~$^{\circ}$C (500~$^{\circ}$F). Unfortunately, two E26 firefighters were located in the flow path between the basement and the doors on the front side of the structure. After a Mayday call, both firefighters were removed from the structure and immediate medical treatment was provided. The two firefighters were transported to the local medical center where the lieutenant was pronounced dead and the fire fighter/paramedic died two days later.},
  pdfkeywords={flow path; fire dynamics; LODD; fire modeling; FDS simulation}
}

@@

The fire originated in the basement, and the interior stairwell acted as a chimney for hot gases in the basement to flow towards areas of lower pressure via the interior stairwell towards the front of the structure. After the rear basement windows began to fail, the combustion gases at elevated temperature and pressure in the basement flowed upwards towards lower pressure areas via the stairwell. The temperature of the gases in the interior stairwell was estimated to be in excess of 700~$^{\circ}$C (1300~$^{\circ}$F), which exceeds the Class~III exposure temperature of 260~$^{\circ}$C (500~$^{\circ}$F). Unfortunately, two E26 firefighters were located in the flow path between the basement and the doors on the front side of the structure. After a Mayday call, both firefighters were removed from the structure and immediate medical treatment was provided. The two firefighters were transported to the local medical center where the lieutenant was pronounced dead and the fire fighter/paramedic died two days later.

@@

Figure~\ref{fig:smv_flow_path_top_1} shows a top view of the temperature contours at a height of 0.3~m (1~ft) above the floor on the first floor, 2~s before the first rear basement window failed. At this time, the temperature of the gases in the interior stairwell were in excess of 200~$^{\circ}$C (400~$^{\circ}$F).

\begin{figure}[!ht]
\includegraphics[width=4.3in]{../Figures/SMV_Top_Temp_340_s}
\input{../Figures/colorbar_metric_temp}
\input{../Figures/colorbar_std_temp}
\caption[Simulated temperature contours on the first floor before basement window failure.]
{Top view of simulated temperature contours on the first floor, 0.3~m (1~ft) above the floor, 2~s before the first rear basement window failure. The orange markers indicate the locations of the downed firefighters.}
\label{fig:smv_flow_path_top_1}
\end{figure}

Figure~\ref{fig:smv_flow_path_top_2} shows a top view of the temperature contours at a height of 0.3~m (1~ft) above the floor on the first floor, 60~s and 120~s after the first rear basement window failed. At a time of 60~s after the first rear basement window failed, the temperature of the gases in the interior stairwell were in excess of 540~$^{\circ}$C (1000~$^{\circ}$F), which exceeds the Class~III exposure temperature of 260~$^{\circ}$C (500~$^{\circ}$F). As more rear basement windows failed and provided ambient air to the growing basement fire, the temperatures in the interior stairwell continued to rise. At a time of 120~s after the first rear basement window failed, the temperature of the gases in the interior stairwell were in excess of 700~$^{\circ}$C (1300~$^{\circ}$F), and the temperatures near the door leading from the interior stairwell into the first floor hallway were in excess of 200~$^{\circ}$C (400~$^{\circ}$F) at a height of 0.3~m (1~ft) above the floor.

@@

\begin{figure}[!ht]
\includegraphics[width=4.3in]{../Figures/SMV_Top_Temp_402_s}
\input{../Figures/colorbar_metric_temp}
\input{../Figures/colorbar_std_temp}
\includegraphics[width=4.3in]{../Figures/SMV_Top_Temp_462_s}
\input{../Figures/colorbar_metric_temp}
\input{../Figures/colorbar_std_temp}
\caption[Simulated temperature contours on the first floor after basement window failure.]
{Top view of simulated temperature contours on the first floor, 0.3~m (1~ft) above the floor, 60~s (top) and 120~s (bottom) after the first rear basement window failure.}
\label{fig:smv_flow_path_top_2}
\end{figure}

Based on these results and the simulation results shown in Sections~\ref{sec:velocity} and \ref{sec:temperature}, after the rear basement windows failed, the simulated flow path temperatures in the interior stairwell were in excess of 700~$^{\circ}$C (1300~$^{\circ}$F) and the flow velocities were approximately 9~m/s (20~mph). The conditions in the interior stairwell changed from tenable to high-hazard very rapidly following the rear basement window failure. Within the structure, the hot gases and smoke were moving along the flow path from the basement towards the doors located on the front side of the structure via the interior stairwell. The simulated temperatures are consistent with the post-incident conditions that were documented in the interior stairwell.

@@

The fire originated in the basement, and the interior stairwell acted as a chimney for hot gases in the basement to flow towards areas of lower pressure via the interior stairwell towards the front of the structure. After the rear basement windows failed, a flow path was established between the basement living room area and the doors located on the front side of the structure (the front door and the garage door). The rear basement window failures resulted in a rapid change in the conditions within the flow path. After the rear basement windows began to fail, the combustion gases at elevated temperature and pressure in the basement flowed upwards towards lower pressure areas via the stairwell. In the interior stairwell, the flow velocities were approximately 9~m/s (20~mph) and the temperature of the gases was estimated to be in excess of 700~$^{\circ}$C (1300~$^{\circ}$F), which exceeds the Class~III exposure temperature of 260~$^{\circ}$C (500~$^{\circ}$F).
